### 2016-06-07 paper.tex
Tighten prose to get down to seven pages

We propose that \totan{} is one of very few published games comprising a genre
of \textit{culture-narration games}. Recognizing that the term ``genre''
can be interpreted in multiple ways, we adopt the perspective established
by \citet{Arsenault2009},
that a genre is ``the codified usage of particular mechanics
and game design patterns to express a range of intended play-experiences.''
We define this genre, then, as having the following characteristics.

@@

\item \textit{Reward players for decisions that reflects
    cultural understanding.}  The knowledge that a player brings to
  bear on decisions is not just knowledge of in-game systems. 
  A player can predict that choosing to fight a Powerful Ifrit 
 without the Weapon Use skill will not end well,
  but this prediction is not made by having memorized the 
  \textit{Book of Tales}.
  Rather, this prediction draws upon a cultural understanding of power,
  spirits, violence, and much more.
  Note that the game rewards player decisions without resorting to
  clear-cut ``right'' and ``wrong'' (``moral/immoral'',
  ``light side/dark side'', ``paragon/renegade'') choices.  There
  remain unexpected twists: perhaps fighting the Powerful Ifrit without
  combat skills makes him respect you and grant a boon. 
  While the world
  remains consistent, it also embraces the potential for surprising
  results.

@@

For contrast, we also consider \textit{Above and Below}~\citep{Laukat2015}, a
worker-management game of city-building.  
%\textit{Above and Below} is primarily about building an
%economic engine that generates victory points faster than your
%opponents within a fixed number of rounds,
Underground exploration includes  player narration
from an encounter book as in \totan{} and \smersh{}, which at first
glance may appear to categorize it as a culture-narration game.
However, in \textit{Above and Below}, reaction choices are thin
wrappers around risk-and-reward balance:
an option is essentially a gamble on being able to roll better
results on the dice, with higher numbers always having better results.
The feedback is primarily in terms of formal rewards, with few
encounters providing any narrative feedback, as illustrated in
the representative encounter of Figure~\ref{fig:encounters}.
Hence, although \textit{Above and Below} includes an encounter book
that players read to each other, it does not satisfy the characteristics
of the culture-narration game genre.

@@

We have discussed how \totan{} uses cultural understanding, not only on a
small scale as a way to reward players, but also as a learning outcome of
the game. In order for the game to reward players for their cultural
understanding, cultures must be presented such that they can be understood
by the player. In other words, the game needs to foster a
sense of \textit{cultural empathy}, defining ``empathy'' as 
the ``state of mind in which someone shares the feelings or outlook of
another, sometimes prompted by imaginative exercises 'stepping into
someone's shoes'''~\citep[p.242]{Honderich2005}
Therefore, cultural
empathy, loosely formed, is having empathy for others who are from
different cultures, which allows for better understanding of that
culture.

@@

Our team therefore chose monsters as a key dramatic element for our 
design experiments. They were not only enticing to our target
audience---first observed by our community partner and later proven
through play testing---but Cohen's theory also 
suggests that they also provide an excellent passage for
developing empathy. Monsters represent
various cultures' sameness, in base forms of fear, yet difference, in
the ways that fear becomes specific and embodied. By engendering an
understanding of both difference and sameness between cultures, they
encourage individuals to try to see something from someone else's
viewpoint.

@@

Physical games such as \totan{} and \smersh{} are expensive to produce
and, for many, intimidating to encounter. \totan{}'s 19-page
instruction booklet may not faze a board game enthusiast, but we have
observed that others are intimidated by the large rulebook, heavy box,
and many components included in the game. Combined with the duration
of play and complexity of rules, such games seem inappropriate for use
in almost all formal educational settings such as conventional
schools.  However, we have found that the combination of technology
enhancement, grade-level writing, and shorter duration circumvents
these problems.  Digital implementations all but eliminate
distribution costs, and with clever game design, children can play the
game in the relatively chaotic environment of an informal after-school
program. With curriculum aides and improved scaffolding, such as
post-game debriefing~\citep{Nicholson2012}
or reflections~\citep{Hickey2013},
we are certain the game could be easily
integrated into school environments, particularly those with
ready access to computers.

>>>>

We propose that \totan{} is one of very few published games comprising
a genre of \textit{culture-narration games}.  This follows the
definition for ``genre'' given by \citet{Arsenault2009}---that it is
``the codified usage of particular mechanics and game design patterns
to express a range of intended play-experiences.''  We define
\textit{culture-narration games}, then, as having the following
characteristics:

@@

\item \textit{Reward players for decisions that reflects
    cultural understanding.}  The knowledge that a player uses in the game
  is not just knowledge of in-game systems. 
  For example, it is not by memorizing the \textit{Book of Tales}
  that leads a player to predict that fighting a Powerful Ifrit without
 the Weapon Use skill will not end well.
  Rather, this prediction draws upon a cultural understanding of power,
  spirits, violence, and much more.
  The game rewards player decisions without resorting to
  binary``right'' and ``wrong'' (``moral/immoral'',
  ``light side/dark side'', ``paragon/renegade'') choices.  There
  remain unexpected twists: perhaps fighting the Powerful Ifrit without
  combat skills makes him respect you and grant a boon. 
  While the world
  remains consistent, it also has the potential for surprising
  results.

@@

For contrast, consider \textit{Above and Below}~\citep{Laukat2015}, a
worker-management game of city-building.  
%\textit{Above and Below} is primarily about building an
%economic engine that generates victory points faster than your
%opponents within a fixed number of rounds,
Underground exploration includes player narration
from an encounter book, which at first
glance may appear to categorize it as a culture-narration game.
However, in \textit{Above and Below}, reaction choices are thin
wrappers around probabilistic risk:
an option is essentially a gamble on being able to roll better
results on the dice, with higher numbers always having better results.
The feedback is primarily in terms of formal rewards, with few
encounters providing any narrative feedback, as illustrated in
the representative encounter of Figure~\ref{fig:encounters}.
Hence, although \textit{Above and Below} includes an encounter book
that players read to each other, it does not satisfy the characteristics
of the culture-narration game genre.

@@

\totan{} uses cultural understanding, not only on a
small scale as a way to reward players, but also as a learning outcome of
the game. In order for the game to reward players for their cultural
understanding, cultures must be presented such that they can be understood
by the player. In other words, the game needs to foster a
sense of \textit{cultural empathy}, defining ``empathy'' as 
the ``state of mind in which someone shares the feelings or outlook of
another, sometimes prompted by imaginative exercises 'stepping into
someone's shoes'''~\citep[p.242]{Honderich2005}
Therefore, cultural
empathy, loosely formed, is having empathy for others who are from
different cultures, which allows for better understanding of that
culture.

@@

Our team therefore chose monsters as a key dramatic element for our 
design experiments. They were not only enticing to our target
audience---first observed by our community partner and later proven
through play testing---but Cohen's theory also 
suggests that they provide an affordance for
developing empathy. Monsters represent
various cultures' sameness, in base forms of fear, yet difference, in
the ways that fear becomes specific and embodied. By engendering an
understanding of both difference and sameness between cultures, they
encourage individuals to try to see something from someone else's
viewpoint.

@@

\totan{}'s 19-page
instruction booklet may not faze a games enthusiast, but 
many are intimidated by the large rulebook, heavy box, and 
number of components.
Combined with the duration
of play, complexity of rules, and relatively high price,
such games are inappropriate for use
in almost all formal educational settings such as conventional
schools.  However, the combination of technology
enhancement, grade-level writing, and shorter duration circumvents
these problems.  Digital implementations all but eliminate
distribution costs, and with clever game design, children can play the
game in the relatively chaotic environment of an informal after-school
program. With curriculum aides and improved scaffolding, such as
post-game debriefing~\citep{Nicholson2012}
or reflections~\citep{Hickey2013},
we are certain the game could be easily
integrated into school environments, particularly those with
ready access to computers.
